---
title: "GISDK Code Reference"
output: html_document
---

GISDK is the scripting language used by TransCAD. In the Tahoe model it provides a user interface for the user and also ties together various sub-modules that are part of the Tahoe Model. This page documents the main GISDK code, [TahoeModelToolbox_rev.rsc](../reference/code/TahoeModelToolbox_rev.rsc) and describes the various functions/methods that are present in the GISDK scripts. In GISDK functions/methods are called  "Macros" and that terminology will be used for the remainder of this document.

The most important Macros are documented below.  For each Macro, the name, the function performed and the output files created will be documented.   

# Model Run Macro Flowchart

### High level flow
```{r,echo=F,,out.height=200}
library(DiagrammeR)
mermaid("
graph LR
A[PreModelRunner]-->B[SaveLinkData]
B-->C{Feedback Loop}
C-->D[CreateTripMatrices]
D-->E[TransitAssignment]
")
```

### PreModelRunner steps

```{r,echo=F,out.height=500}
mermaid("
graph TD
B[UpdateFFSpeeds]-->C[CreateDriveNetwork]
C-->D[CopyBaseTripTables]
D-->E[CreateTripMatrices]
E-->F[TrafficAssignment]
F-->G[BandsRun]
G-->H[TransitSkimmer]
H-->I[JavaPreModelCode]
")
```

### Inside the feedback loop
```{r,echo=F,out.height=500}
mermaid("
graph TD
L[ClearLastFile]-->M[RunJavaModel]
M-->N[CreateTripMatrices]
N-->O[TrafficAssignment]
O-->P[MSAAssignmentResults]
P-->Q[RunTransitSkims]
Q-->R[SaveLinkData]
R-->S[TripSummarizer]
")
```

# Macro Reference

---------------------------------------------------------------------------------------
Macro                Macro                                          Output files from
Name                 Description                                    the Macro
-------------------- ---------------------------------------------- -------------------
PreModelRunner       This is the first Macro that is run      Refer to the outputs 
                     when the "Run Model" button is clicked.        of Macros that are
                     It is outside the global feedback loop and     called by this Macro  
                     runs several other Macros listed below. 
                     `UpdateFFSpeeds`<br/>
                     `CreateDriveNetwork`<br/>
                     `CopyBaseTripTables`<br/>
                     `CreateTripMatrices`<br/>
                     `TrafficAssignment`<br/>
                     `BandsRun`<br/>
                     `TransitSkimmer`<br/>
                     `JavaPreModelCode`<br/>
                     
                     
UpdateFFSpeeds       Updates the link free flow speed in the 
                     input network based on a specific formula.     FullStreets.bin
                     
CreateDriveNetwork   Creates the Tahoe Drive Network                Tahoe_Drive_Network.net 
                     
CopyBaseTripTables   Creates a copy of the base trip table so that  Trips_#PRD#.csv for
                     it can be used to create the base trip         all time periods
                     matrices
                     
CreateTripMatrices   Converts each of the four CSV trip tables      Trips_#PRD#.mtx for
                     created in the previous step into TransCAD     all time periods
                     trip matrices in the .mtx format
                     
TrafficAssignment    
                     
BandsRun             This macro creates long and short walk bands    WalkAccess_#SEASON#.csv
                      for the Tahoe region (by Season)
                     
TransitSkimmer       Creates transit skims by skimming the 
                      transit network. Also calls the macro
                      JavaSkimConverter, which converts the 
                      transit skims in TransCAD matrix format into
                      *.zmx formats - so that it can be read later
                      by the Java program.
                      
                     
JavaPreModelCode     
                     
SaveLinkData         This Macro saves the link data from the         Fullstreets.bin  
                     network file to fullstreets.bin with the
                     feedback loop iteration suffixed to the
                     file
                     
ClearLastFile        
                     
RunJavaModel         Runs the main Java model, the demand model. 
                      It runs the followings demand sub-modules:
                      resident, external workers and visitors.
                     
MSAAssignmentResults 

RunTransitSkims                   

TripSummarizer        Calls the TahoeResultsSummarizer (an external
                        Python executable) to create summary reports
                        
TransitAssignment     Runs transit assignment

----------------------------------------------------------------------------------




