---
title: "Dynamic Validation: Recreational Attraction Changes"
output: 
  html_document
---


```{r include=FALSE,messages=F}
library(tidyverse)
library(readxl)
library(kableExtra)
library(rgdal)
library(sf)
library(rgeos)
library(htmltools)
library(htmlwidgets)
library(leaflet)
library(DT)

convAttrToChar <- function(spdf){
    spdf@data <- spdf@data %>% mutate_if(is.factor,as.character)
    return(spdf)
}

mpal2 <- colorNumeric(c('#969696','#0570b0'), domain = 1:2) # For coding centroid connectors vs. other links

```



```{r include=F,messages=F}
# Base Year Model
scen <- '37_BY'
FullStreets <- read_csv(paste0('Runs/',scen,'/post/FullStreets.csv'))
FullStreets <- FullStreets %>% mutate(DAILYVOLUME=round(DAILYVOLUME))
streetNet <- st_read(paste0('Runs/',scen,'/post/Streets_offset.shp')) %>% as('Spatial') %>% convAttrToChar()
nodesLayr <- st_read(paste0('Runs/',scen,'/post/Nodes_offset.shp')) %>% as('Spatial') %>% convAttrToChar()
nodesLayr@data$TAZ[is.na(nodesLayr@data$TAZ)] <- 0
taz289 <- nodesLayr[nodesLayr$TAZ==289,]

streetNet@data <- streetNet@data %>% select(ID) %>% left_join(FullStreets %>% select(ID,AB_FC,AB_SPEED,DAILYVOLUME),by=c('ID'='ID')) 

streetNet@data <- streetNet@data %>% mutate(tazLabel = paste0('<i>ID: </i>',ID,'<br>',
                                'Facility Type: ',AB_FC,'<br>',
                                'Speed =',AB_SPEED,'<br>',
                                'Model Volume: ',DAILYVOLUME,'<br>'))
streetNet_FC129 <- streetNet[streetNet$AB_FC %in% c(1,2,9),]
streeCen = (data.frame(gCentroid(streetNet_FC129,byid=TRUE)))
streeCen <- streeCen %>% mutate(DAILYVOLUME=streetNet_FC129$DAILYVOLUME)

mm_base <-leaflet(options = leafletOptions(preferCanvas = TRUE)) %>% 
  addProviderTiles(providers$Esri.WorldGrayCanvas, options = providerTileOptions(updateWhenZooming = FALSE,updateWhenIdle = TRUE),group='WorldGray') %>% 
  addProviderTiles(providers$Esri.WorldStreetMap, options = providerTileOptions(updateWhenZooming = FALSE,updateWhenIdle = TRUE),group='WorldMap') %>% 
  addPolylines(data=streetNet,group='All Links',color=~mpal2(ifelse(AB_FC==9,2,1)),weight = 2,
               label=lapply(streetNet@data$tazLabel, function(x) HTML(x)),
               labelOptions = labelOptions(noHide = F,offset = c(10,-20))) %>% 
  addCircleMarkers(lng = streeCen$x, lat = streeCen$y,label = as.character((streeCen$DAILYVOLUME)),group='Model Volume',opacity=0,radius = 1,
                   labelOptions = lapply(1:nrow(streeCen), function(x) {
                     labelOptions(opacity=1, noHide = T,textOnly = TRUE,textsize='10px',direction = 'auto', offset=c(0,0)) 
                   })) %>%   
  addControl(HTML('<font size="5">Base Year</font>'), position = "bottomleft") %>% 
  addCircleMarkers(data=taz289,opacity=0,group='TAZ 289') %>% 
  addLayersControl(baseGroups = c('WorldGray','WorldMap'),overlayGroups = c('All Links','Model Volume','TAZ 289'),options = layersControlOptions(collapsed = F)  ) %>% 
  hideGroup(c('Model Volume'))

```

## Reduce King's Beach Recreational Attractiveness

In this dynamic validation test, the recreational attractiveness of King's Beach TAZ is decreased by 50%. In the base year the recreational attractiveness of King's Beach TAZ had a value of 10,656. This was decreased to 5,328 in the dynamic validation test. King's beach is located in a TAZ with ID 289 as shown in the maps below using a blue circle. 

The model was run with this update and the resulting highway loading is shown in the map below on the right. The base year highway loaded result is shown on the left for comparison.

```{r include=F,messages=F}
scen <- '38_Rec'
FullStreets <- read_csv(paste0('Runs/',scen,'/post/FullStreets.csv'))
FullStreets <- FullStreets %>% mutate(DAILYVOLUME=round(DAILYVOLUME))
streetNet <- st_read(paste0('Runs/',scen,'/post/Streets_offset.shp')) %>% as('Spatial') %>% convAttrToChar()

streetNet@data <- streetNet@data %>% select(ID) %>% left_join(FullStreets %>% select(ID,AB_FC,AB_SPEED,DAILYVOLUME),by=c('ID'='ID')) 

streetNet@data <- streetNet@data %>% mutate(tazLabel = paste0('<i>ID: </i>',ID,'<br>',
                                'Facility Type: ',AB_FC,'<br>',
                                'Speed =',AB_SPEED,'<br>',
                                'Model Volume: ',DAILYVOLUME,'<br>'))
streetNet_FC129 <- streetNet[streetNet$AB_FC %in% c(1,2,9),]
streeCen = (data.frame(gCentroid(streetNet_FC129,byid=TRUE)))
streeCen <- streeCen %>% mutate(DAILYVOLUME=streetNet_FC129$DAILYVOLUME)

mm <-leaflet(options = leafletOptions(preferCanvas = TRUE)) %>% 
  addProviderTiles(providers$Esri.WorldGrayCanvas, options = providerTileOptions(updateWhenZooming = FALSE,updateWhenIdle = TRUE),group='WorldGray') %>% 
  addProviderTiles(providers$Esri.WorldStreetMap, options = providerTileOptions(updateWhenZooming = FALSE,updateWhenIdle = TRUE),group='WorldMap') %>% 
  addPolylines(data=streetNet,group='All Links',color=~mpal2(ifelse(AB_FC==9,2,1)),weight = 2,
               label=lapply(streetNet@data$tazLabel, function(x) HTML(x)),
               labelOptions = labelOptions(noHide = F,offset = c(10,-20))) %>% 
  addCircleMarkers(lng = streeCen$x, lat = streeCen$y,label = as.character((streeCen$DAILYVOLUME)),group='Model Volume',opacity=0,radius = 1,
                   labelOptions = lapply(1:nrow(streeCen), function(x) {
                     labelOptions(opacity=1, noHide = T,textOnly = TRUE,textsize='10px',direction = 'auto', offset=c(0,0)) 
                   })) %>%   
  addControl(HTML('<font size="5">Kings Beach: Less Attractive</font>'), position = "bottomleft") %>% 
  addCircleMarkers(data=taz289,opacity=0,group='TAZ 289') %>% 
  addLayersControl(baseGroups = c('WorldGray','WorldMap'),overlayGroups = c('All Links','Model Volume','TAZ 289'),options = layersControlOptions(collapsed = F)  ) %>% 
  hideGroup(c('Model Volume'))

```

<div style= "float:left;position: relative; top: 10px;">
```{r echo=FALSE, fig.height=4,fig.width=4.7}
mm_base %>%  setView(lng = -120.022618 , lat = 39.235187, zoom = 15) %>% showGroup(c('Model Volume'))
```
</div>


<div style= "float:right;position: relative; top: 10px;">
```{r echo=FALSE, fig.height=4,fig.width=4.7}
mm %>%  setView(lng = -120.022618 , lat = 39.235187, zoom = 15) %>% showGroup(c('Model Volume'))
```
</div>

Key insights from this dynamic testing is summarized below:

+ The number of vehicle trips starting/ending in King's Beach TAZ decreased from 2338 trips to 1504 trips, a reduction of 36%.

+ There is no change in the number of visitors coming into the Lake Tahoe region - since the socio-economic data and occupancy data was not changed in this test. Despite that there were some changes in the system wide trip/VMT statistic. The overall VMT increased only marginally (0.01%); however, there were larger differences when the VMT of different trip types were analyzed. The VMT of internal-internal trips decreased by 0.33% while the VMT of internal-external and external-internal trips increased by 0.13%. One explanation of this change is that residents and overnight visitors, seeing a reduction in recreational opportunities in the Lake Tahoe area, are going outside the region to undertake recreational activities.

