---
title: "Static Validation"
output: 
  html_document
---


```{r include=FALSE,messages=F}

library(tidyverse)
library(readxl)
library(knitr)
library(kableExtra)
library(plotly)

library(rgdal)
library(sf)
library(rgeos)
library(htmltools)
library(leaflet)
library(DT)

knitr::opts_knit$set(knitr.kable.NA = '')
scen <- '53_Reid'

### read in and combine distances skim files ###
### read in and combine distances skim files ###
SummerLateNightDriveDistanceSkim <- read_csv(paste0('../Calibration/Runs/',scen,'/gis/skims/SummerAMPeakDriveDistanceSkim.csv')) 
SummerLateNightDriveDistanceSkim <- setNames(SummerLateNightDriveDistanceSkim,c('Orig','Dest','Ivtt','DistanceInSkim'))

extDistMap <- read_excel('../Calibration/Data/0_VariableMap.xlsx','ExtStationDistances','A1:B15')


SocioEcon <- read_csv(paste0('../Calibration/Runs/',scen,'/zonal/SocioEcon_Summer.csv'))
OvernightVisitorZonalData <- read_csv(paste0('../Calibration/Runs/',scen,'/zonal/OvernightVisitorZonalData_Summer.csv'))
VisitorOccupancyRates <- read_csv(paste0('../Calibration/Runs/',scen,'/zonal/VisitorOccupancyRates_Summer.csv'))
VisitorReportData <- read_csv(paste0('../Calibration/Runs/',scen,'/outputs_summer/VisitorReportData.csv')) %>% 
  mutate(visitorTypeS = case_when(visitorType==1~'overnight visitor',visitorType==2~'day visitor',visitorType==3~'thru visitor'))



extTAZCutOff <- 108
SummerLateNightDriveDistanceSkim <- SummerLateNightDriveDistanceSkim %>% left_join(extDistMap,by=c('Orig'='TAZ')) %>% rename(ed1=ExternalDistance) %>% 
  left_join(extDistMap,by=c('Dest'='TAZ')) %>% rename(ed2=ExternalDistance) %>% replace_na(list(ed1=0,ed2=0)) %>% 
  mutate(ExtDist=ed1+ed2) %>% 
  mutate(DistanceTotal = ifelse(Orig<=extTAZCutOff & Dest<=extTAZCutOff,DistanceInSkim+ExtDist,DistanceInSkim),
         DistanceWithinRegion = DistanceTotal - ExtDist )

### read in and combine distances skim files ###
pm<-read_csv(paste0('../Calibration/Runs/',scen,"/gis/Skims/SummerPMPeakDriveDistanceSkim.csv")) %>%
  mutate(skim=2,id2=paste(`TAZ:1`, TAZ,skim, sep="-")) %>%
  rename(trip_time=`AB_PM_IVTT / BA_PM_IVTT`)
am<-read_csv(paste0('../Calibration/Runs/',scen,"/gis/Skims/SummerAMPeakDriveDistanceSkim.csv")) %>%
  mutate(skim=1,id2=paste(`TAZ:1`, TAZ,skim, sep="-")) %>%
  rename(trip_time=`AB_AM_IVTT / BA_AM_IVTT`)
ln<-read_csv(paste0('../Calibration/Runs/',scen,"/gis/Skims/SummerLateNightDriveDistanceSkim.csv")) %>%
  mutate(skim=4,id2=paste(`TAZ:1`, TAZ, skim, sep="-")) %>%
  rename(trip_time=`AB_LN_IVTT / BA_LN_IVTT`)
md<-read_csv(paste0('../Calibration/Runs/',scen,"/gis/Skims/SummerMiddayDriveDistanceSkim.csv")) %>%
  mutate(skim=3, id2=paste(`TAZ:1`, TAZ, skim, sep="-")) %>%
  rename(trip_time=`AB_MD_IVTT / BA_MD_IVTT`)
dist <- bind_rows(pm,am,ln,md) %>%
  mutate(ext_dist=case_when(`TAZ:1`%in% c(1,2,3,4,5,6,7,10,20,30,40,50,60,70) & 
                              TAZ %in% c(1,2,3,4,5,6,7,10,20,30,40,50,60,70) ~ 0,
                            `TAZ:1` == 1 | TAZ == 1 ~ 25.1,
                            `TAZ:1` == 2 | TAZ == 2 ~ 21.6,
                            `TAZ:1` == 3 | TAZ == 3 ~ 17.2,
                            `TAZ:1` == 4 | TAZ == 4 ~ 16.7,
                            `TAZ:1` == 5 | TAZ == 5 ~ 30.9,
                            `TAZ:1` == 6 | TAZ == 6 ~ 19.0,
                            `TAZ:1` == 7 | TAZ == 7 ~ 15.1,
                            `TAZ:1` == 10 | TAZ == 10 ~ 136,
                            `TAZ:1` == 20 | TAZ == 20 ~ 114,
                            `TAZ:1` == 30 | TAZ == 30 ~ 424,
                            `TAZ:1` == 40 | TAZ == 40 ~ 408,
                            `TAZ:1` == 50 | TAZ == 50 ~ 112,
                            `TAZ:1` == 60 | TAZ == 60 ~ 130,
                            `TAZ:1` == 70 | TAZ == 70 ~ 122                            
  ),
  distance=`Length (Skim)`- ext_dist) %>%
  mutate(distance = ifelse(is.na(distance),`Length (Skim)`,distance),
         time_woExtDist = trip_time - (`Length (Skim)`-distance)/50*50) %>% 
  mutate(
    ext_dist = ifelse(`TAZ:1`%in% c(1,2,3,4,5,6,7,10,20,30,40,50,60,70) & 
                        TAZ %in% c(1,2,3,4,5,6,7,10,20,30,40,50,60,70),
                      ext_dist+(
                        case_when(`TAZ:1` == 1|`TAZ:1` == 10~25.1,
                                  `TAZ:1` == 2|`TAZ:1` == 20~21.6,
                                  `TAZ:1` == 3|`TAZ:1` == 30~17.2,
                                  `TAZ:1` == 4|`TAZ:1` == 40~16.7,
                                  `TAZ:1` == 5|`TAZ:1` == 50~30.9,
                                  `TAZ:1` == 6|`TAZ:1` == 60~19.0,
                                  `TAZ:1` == 7|`TAZ:1` == 70~15.1)
                      ),
                      ext_dist
    ),
    ext_dist = ifelse(`TAZ:1`%in% c(1,2,3,4,5,6,7,10,20,30,40,50,60,70) & 
                        TAZ %in% c(1,2,3,4,5,6,7,10,20,30,40,50,60,70),
                      ext_dist+(
                        case_when(TAZ == 1|TAZ == 10~25.1,
                                  TAZ == 2|TAZ == 20~21.6,
                                  TAZ == 3|TAZ == 30~17.2,
                                  TAZ == 4|TAZ == 40~16.7,
                                  TAZ == 5|TAZ == 50~30.9,
                                  TAZ == 6|TAZ == 60~19.0,
                                  TAZ == 7|TAZ == 70~15.1)
                      ),
                      ext_dist
    )
  ) %>% 
  mutate(
    `Length (Skim)` = ifelse(
      `TAZ:1`%in% c(1,2,3,4,5,6,7,10,20,30,40,50,60,70) & 
        TAZ %in% c(1,2,3,4,5,6,7,10,20,30,40,50,60,70),
      `Length (Skim)`+ext_dist,
      `Length (Skim)`
    ),
    trip_time = ifelse(
      `TAZ:1`%in% c(1,2,3,4,5,6,7,10,20,30,40,50,60,70) & 
        TAZ %in% c(1,2,3,4,5,6,7,10,20,30,40,50,60,70),
      trip_time+ext_dist,
      trip_time
    )
  )

dist <- dist %>% select(`TAZ:1`,TAZ,skim,`Length (Skim)`,distance) %>% 
  rename(Orig=`TAZ:1`,Dest=TAZ,DistanceTotal=`Length (Skim)`,DistanceWithinRegion=distance) %>% 
  mutate(ExtDist = DistanceTotal-DistanceWithinRegion )




trip_file <- read_csv(paste0('../Calibration/Runs/',scen,'/outputs_summer/trip_file.csv'))
PartyArray <-  read_csv(paste0('../Calibration/Runs/',scen,'/outputs_summer/PartyArray_afterThruVisitors.pam'))
extDistMap <- read_excel('../Calibration/Data/0_VariableMap.xlsx','ExtStationDistances','A1:B15')





trip_file <- trip_file %>% 
  left_join(PartyArray %>% select(id,stayType,walkSegment), by=c("partyID"="id")) %>% 
  mutate(partyType=ifelse(partyType=='overnight visitor' & stayType==1,'seasonal',partyType)) %>% 
  mutate(mode = case_when(mode=='drive alone'~'Drive Alone',
                          mode=='shared auto'~'Shared Auto',
                          mode=='visitor shuttle'~'Shuttle',
                          mode=='walk to transit'~'Walk to Transit',
                          mode=='drive to transit'~'Drive to Transit',
                          mode=='non motorized'~'Non-motorized',
                          mode=='school bus'~'School Bus',
  )) %>% 
  mutate(mode=factor(mode,levels=c('Drive Alone', 'Shared Auto', 'Shuttle','Walk to Transit','Drive to Transit','Non-motorized','School Bus')))

trip_file<- trip_file %>% left_join(dist,by=c('startTaz'='Orig','endTaz'='Dest','skim'))

trip_file <- trip_file %>% filter(mode %in% c('Drive Alone','Shared Auto'))

gateSumm1_Target <- read_excel('../Calibration/Data/1_Targets_FromStreetlightData.xlsx','Aggregate_Calculation','A3:I11')
gateSumm1_Target <- gateSumm1_Target %>% select(-`EGate`,-ExternalGate)
gateSumm1_Target <- gateSumm1_Target %>% mutate_at(2:7,round,0)



systemLevelTarget <- read_excel('../Calibration/Data/1_Targets_FromStreetlightData.xlsx','VMT_n_Trips','B2:E9')

trip_file <- trip_file %>% mutate(TripCat = ifelse(startTaz<extTAZCutOff & endTaz<extTAZCutOff,'EE Trips',
                                                   ifelse(startTaz<extTAZCutOff | endTaz<extTAZCutOff, 'IE Trips','II Trips'
                                                   )))

EE_Target <- read_excel('../Calibration/Data/1_Targets_FromStreetlightData.xlsx','EE_Trips','A1:I9')
EE_Target[is.na(EE_Target)] <- 0


extTripsTarget <- read_excel('../Calibration/Data/1_Targets_FromStreetlightData.xlsx','Long_vs_Short_Ext','B2:D10')
extTripsVMTTarget <- read_excel('../Calibration/Data/1_Targets_FromStreetlightData.xlsx','Long_vs_Short_Ext','B12:D20')



convAttrToChar <- function(spdf){
    spdf@data <- spdf@data %>% mutate_if(is.factor,as.character)
    return(spdf)
}

FullStreets <- read_csv(paste0('../Calibration/Runs/',scen,'/post/FullStreets.csv'))


SummerAMPeakboardAlight <- read_csv(paste0('../Calibration/Runs/',scen,'/post/SummerAMPeakboardAlight.csv'))
SummerLateNightboardAlight <- read_csv(paste0('../Calibration/Runs/',scen,'/post/SummerLateNightboardAlight.csv'))
SummerMiddayboardAlight <- read_csv(paste0('../Calibration/Runs/',scen,'/post/SummerMiddayboardAlight.csv'))
SummerPMPeakboardAlight <- read_csv(paste0('../Calibration/Runs/',scen,'/post/SummerPMPeakboardAlight.csv'))


TripMatrixModeMap <- read_excel('../Calibration/Data/0_VariableMap.xlsx','TripMatrixModeMap','A1:B8')

highwayCounts <- read_excel('../Calibration/Data/3_CountsAndBoardings.xlsx','HighwayCounts','A1:F35') %>% select(-SL_vol,-diff,-AB_LANES)

boardingTargets <- read_excel('../Calibration/Data/3_CountsAndBoardings.xlsx','TransitBoardsings','B2:D15')
transitRouteMap <- read_excel('../Calibration/Data/3_CountsAndBoardings.xlsx','TransitBoardsings','G2:H28')


streetNet <- st_read(paste0('../Calibration/Runs/',scen,'/post/Streets_offset.shp')) %>% as('Spatial') %>% convAttrToChar()

Tahoe_TransitR <- read_csv(paste0('../Calibration/Runs/',scen,'/post/Tahoe_TransitR.csv'))


```


This report presents the summary of the model results after calibration of activity-based model sub-models and validation of highway assignment and transit assignment procedures. The first sub-section *External Station Validation* compares the model forecasted trips passing through the external stations against the corresponding targets prepared in an earlier report (ExternalStationComposition.html). The *Highway Validation* section provides the summary of highway validation by comparing model assigned volume against observed roadway counts. The *Transit Validation* sections compares the boardings predicted by model against the observed boardings. The last section looks at the vehicle miles travelled (VMT) by all the trips; within, into and out of the Lake Tahoe modeling region.

## External Station Validation

The external stations trip composition calibration results are shown in the table below. The creation of the target trip numbers is described in detail in the report ExternalStationComposition.html. The calibration of work trips; for both residents and external workers, has also been detailed in the report ExternalWorkerCalibration.html. 


```{r, include=T,messages=F,echo=F}

extTrips <- trip_file %>% filter((startTaz!=0 & startTaz <= extTAZCutOff)|(endTaz!=0 & endTaz <= extTAZCutOff)) %>% 
  mutate(tripCat = ifelse(partyType=='thru visitor','Through trips',
                          ifelse(partyType=='resident' & tripType=='Work'|partyType=='external worker' ,'Work trips',
                                 ifelse(partyType=='resident' & tripType=='School','School trips',
                                        ifelse(partyType=='resident','Resident DayTrips',
                                               ifelse(partyType=='overnight visitor','Overnight Visitor DayTrips','Day Visitors')))
                                 )),
         extTaz = ifelse(startTaz<=extTAZCutOff & startTaz!=0,startTaz,endTaz),
         extGate = ifelse(extTaz>9,extTaz/10,extTaz))

gateSumm1 <- extTrips %>% group_by(tripCat,extGate) %>% summarise(numtrips=n()) %>% spread(tripCat,numtrips)
gateSumm2 <-  extTrips %>% filter(tripCat=='Through trips') %>% mutate(extTaz=endTaz,extGate = ifelse(extTaz>9,extTaz/10,extTaz)) %>% 
  group_by(tripCat,extGate) %>% summarise(numtrips=n()) %>% spread(tripCat,numtrips)

gateSumm1$`Through trips` <- gateSumm1$`Through trips`+gateSumm2$`Through trips`  

gateSumm1 <- gateSumm1 %>% select(extGate,`Work trips`,`Through trips`,`Resident DayTrips`,`Overnight Visitor DayTrips`,`Day Visitors`) %>% rename(`External Gate`=extGate)
gateSumm1 <- gateSumm1 %>% mutate(`Total Trips` =  `Through trips` +`Work trips`+ `Overnight Visitor DayTrips`+ `Resident DayTrips` +`Day Visitors`)
  
gateSumm1 <- rbind.data.frame(gateSumm1,gateSumm1 %>% summarise_if(is.numeric,sum) %>% 
                                                       mutate(`External Gate`='Total'))
gateSumm1$`External Gate` <- gateSumm1_Target$`External Gate`

gateSumm1_Target %>% kable(caption='Table 1: External trips by type and gate: Targets',format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

gateSumm1 %>% kable(caption='Table 2: External trips by type and gate: Model',format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 

```

The table above shows that the match between the target trips and modeled trips by external station and market segment is close. The details on the sub-model that was calibrated for each of the market segment is presented below.

+ Through trips - through visitor population synthesis is controlled by the external station specific parameter *percentThru* in the parameter file *DayVisitorZonalData_Summer.csv*. This parameter was calibrated to match the number of observed through trips at the origin zones. Through visitor destinations are modeled in the Through Visitor Destination choice model (ThruVisitorDC.xls). The constants in this model were updated to calibrate the through visitor destinations. 

+ Resident Day Trips	- Resident non-work tour destination choices are specified in the parameter file *Dest_Choice.xls*. The destination choice models are segmented by tour purposes and all the tour purposes were updated to match the target number of resident non-work trips that cross an external station. 

+ Overnight Visitor Day Trips - the overnight visitor destination choice model was calibrated to match the targets for overnight visitor day trips. This model is segmented by tour purposes and is specified in the parameter file *VisitorDC.xls*.

+ Day Visitor Trips - Day visitor population synthesis is controlled by the external station specific parameter *overnight2day* in the parameter file *DayVisitorZonalData_Summer.csv*.  This external station specific parameter was calibrated to match the observed day visitor trips to/from that external zone.

## Highway Validation

```{r,echo=F,eval=T}

counthighwaySTL <- read_excel('../Calibration/Data/3_CountsAndBoardings.xlsx','HighwayCounts','A1:F35') %>% select(-diff,-AB_LANES) %>% 
  filter(Source=='DOT Count')

mdl <- lm(count_vol ~ 0 + SL_vol, data = counthighwaySTL)
rsq <- signif(summary(mdl)$adj.r.squared, 2)

```


```{r,echo=F,eval=T}
gg <- ggplot(counthighwaySTL, aes(SL_vol, count_vol))+
    geom_point() + geom_smooth(method = "lm", se = F) +
     # scale_x_continuous(limits = c(0, 40000)) +
     # scale_y_continuous(limits = c(0, 40000)) +
    labs(x = "Streetlight Data Count", y="DOT Count", title = "Streetlight Data Count vs. DOT Count")
# pp <- gg %>% ggplotly() 
# pp <- pp %>% layout(annotations =list(x=max(tazWorkerCompare$SEDataEmployment)*0.66,y=max(tazWorkerCompare$residentWorkers)*0.33,text=paste0('R-squared = ',rsq),showarrow=F))

```

<div style= "float:right;position: relative; top: 10px;">
```{r pressure, echo=FALSE, fig.height=4,fig.width=5}
gg
```
</div>

In the highway assignment the origin-destination trip demand matrix created from the demand models is assigned (loaded) on to the network. The primary output from the highway assignment step is the traffic flow on each roadway link, commonly referred to as the model volume. This section presents the comparison of the model volume against the roadway count obtained from the count databases of the transportation agencies - Nevada Department of Transportation and California Department of Transportation’s Caltrans Performance Measurement System (PeMS) Data. As a consistency check on the roadway count data and the StreetLight data, the two sources were compared to each other. The initial benefit of this comparison was that it helped unearth an error made in summarizing the count data during the initial compilation of the roadway count data. Using the correctly summarized roadway count data, a comparison was made against StreeLlight counts by computing the R2, across all the links that had a roadway count information (see plot to the right).  An R2 value of 0.97 indicates a strong match between the roadway count data (used to validate highway assignment) and Streetlight Data (used to validate external station traffic and composition); thus providing confidence in the count data and the StreetLight data. Only five, out of the twenty-five links that had roadway counts, had a Streetlight Data count deviation of 25% or more  from the roadway count. The counts obtained from transportation agencies did not cover some of the regions in the modeling area, especially west shore and east shore. To fill this missing gap, the count from StreetLight Data was used. The link ID for the roadway links where StreetLight Data was used are the following: 145,3068,12135,14493,622,4881,4077,14579,14499,2439. Table 3 presents the overall summary and Table 4 provides the link level detail.


```{r, include=T,messages=F,echo=F}

streetNet2 <- streetNet
linkSumm <- highwayCounts %>% left_join(FullStreets %>% select(ID,DAILYVOLUME),by=c('model_link'='ID')) %>% 
  mutate_if(is.numeric,round)

linkSumm <- linkSumm %>% mutate(`Model/Count`=round(DAILYVOLUME/count_vol,2),
                                `Model Devation` = round(abs(1-`Model/Count`),2),
                                `Max Deviation`=
                                  ifelse(count_vol<3000,0.6,
                                         ifelse(count_vol<6000,0.44,
                                                ifelse(count_vol<13000,0.32,
                                                       ifelse(count_vol<27000,0.26,0.24)))),
                                `Within Deviation` = ifelse(`Model Devation`<=`Max Deviation`,'Yes','No'))
options(knitr.kable.NA = '')

rmse <- linkSumm %>% mutate(sqdiff = (count_vol-DAILYVOLUME)^2) %>% summarise(rmse=sqrt(mean(sqdiff)))
pctRMSE <- paste0(round(rmse/(sum(linkSumm$count_vol)/nrow(linkSumm))*100,1)$rmse,'%')
corVal <- round(cor(linkSumm$count_vol,linkSumm$DAILYVOLUME),2)
pctWithinMaxDev <- paste0(round(mean(linkSumm$`Within Deviation`=='Yes')*100,1),'%')

dfStat <- data.frame(Statistic=c('% RMSE Overall','Correlation Coefficient','Percent of Links Within Allowable Deviation'),
                     Model=c(pctRMSE,corVal,pctWithinMaxDev),
                     Threshold=c('Below 40%', 'At Least 0.88','At least 75%'))
dfStat %>% kable(caption='Table 3 Highway validation statistic') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

streetNet2@data <- streetNet2@data %>% left_join(linkSumm %>% select(model_link,`Within Deviation`) %>% rename(ID=model_link),by='ID')

linkSumm <- linkSumm %>% mutate(model_link=as.character(model_link))

tempRowSum <- linkSumm %>% ungroup() %>% summarise_if(is.numeric,sum) %>% 
  mutate(model_link='Total',`Within Deviation`="",
         `Model/Count`=round(DAILYVOLUME/count_vol,2),`Model Devation`=NA,`Max Deviation`=NA,Source=NA)
linkSumm <- rbind.data.frame(linkSumm,tempRowSum) %>% rename(`Link Id`=model_link,
                                                             `Count Volume`=count_vol,`Model Volume`=DAILYVOLUME,`Count Source` = Source)
linkSumm <- linkSumm %>% select(1,2,4,5,6,7,8,3)

```

In Table 4 the first four columns are self-explanatory. The column *Model Deviation* is defined as the *Abs(1-Model/Count)*. *Max Deviation* is based on criterion developed by *2010 RTP Guidelines*. It defined maximum deviation that is acceptable for each link based on the observed count - the larger the observed count, the smaller the acceptable deviation. The last column compares the Model Deviation and the Max Deviation and notes whether the link satisfies the max deviation criterion or not. 

```{r, include=T,messages=F,echo=F}

DT::datatable(linkSumm,caption = 'Table 4 Observed Link Count vs. Daily Volume',
          options = list(pageLength = 12, info = FALSE)) 


```




```{r, include=T,messages=F,echo=F}

# %>% kable(caption = 'Table 2.2 Link Count versus Daily Volume')   %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 

streetNet2@data <- streetNet2@data %>% select(ID,`Within Deviation`) %>% left_join(FullStreets %>% select(ID,AB_FC,AB_SPEED,DAILYVOLUME),by=c('ID'='ID'))  
streetNet2@data <- streetNet2@data %>% left_join(highwayCounts,by=c('ID'='model_link')) %>% 
  mutate_if(is.numeric,round) %>% 
  mutate(
    AbsDiff = ifelse(!is.na(count_vol),abs(DAILYVOLUME-count_vol),NA),
    PctDiff = ifelse(!is.na(count_vol),round((DAILYVOLUME-count_vol)/count_vol*100,0),NA))

streetNet2@data <- streetNet2@data %>% mutate(tazLabel = paste0('<i>ID: </i>',ID,'<br>',
                                'Facility Type: ',AB_FC,'<br>',
                                'Speed =',AB_SPEED,'<br>',
                                'Model Volume: ',DAILYVOLUME,'<br>',
                                'Count Volume: ',count_vol,'<br>',
                                'Abs Difference: ',AbsDiff,'<br>',
                                'Pct Diff: ',PctDiff,'<br>'))

streetNet2_wCounts <- streetNet2[!is.na(streetNet2$count_vol),]

streetNet2_FC12 <- streetNet2[streetNet2$AB_FC %in% c(1,2),]

streeCen = (data.frame(gCentroid(streetNet2_FC12,byid=TRUE)))
streeCen <- streeCen %>% mutate(DAILYVOLUME=streetNet2_FC12$DAILYVOLUME)

streetNet2_wCounts@data <- streetNet2_wCounts@data %>% 
  mutate(thickCat = case_when(AbsDiff <  2500 & AbsDiff>=     0~4,
                              AbsDiff <  5000 & AbsDiff>=  2500~5,
                                                AbsDiff>=  5000~6,
                              ),
         colorCat = case_when(PctDiff < -50                ~1,
                              PctDiff < -25 & AbsDiff>= -50~2,
                              PctDiff <  0  & AbsDiff>= -25~3,
                              PctDiff <  25 & AbsDiff>=   0~4,
                              PctDiff <  50 & AbsDiff>=  25~5,
                                              AbsDiff>=  50~6,
                              ),
         )

streetNet2_wCountsCen = (data.frame(gCentroid(streetNet2_wCounts,byid=TRUE)))
streetNet2_wCountsCen <- streetNet2_wCountsCen %>% mutate(ID=streetNet2_wCounts$ID,withinDevYes =1*(streetNet2_wCounts$`Within Deviation`=='Yes' ))


pal <- colorNumeric("RdYlBu", domain = 1:6)
mpal <- colorNumeric(c('#b30000','#d7301f','#ef6548','#66c2a4','#238b45','#006d2c'), domain = 1:6)
mpal2 <- colorNumeric(c('#969696','#0570b0'), domain = 1:2) # For coding centroid connectors vs. other links

```

```{r, include=T,messages=F,echo=F, out.height=800,out.width=800}

mm <-leaflet(options = leafletOptions(preferCanvas = TRUE)) %>% 
    addProviderTiles(providers$Esri.WorldGrayCanvas, options = providerTileOptions(
        updateWhenZooming = FALSE,      # map won't update tiles until zoom is done
        updateWhenIdle = TRUE           # map won't load new tiles when panning
    ),group='WorldGray') %>% 
    addProviderTiles(providers$Esri.WorldStreetMap, options = providerTileOptions(
        updateWhenZooming = FALSE,      # map won't update tiles until zoom is done
        updateWhenIdle = TRUE           # map won't load new tiles when panning
    ),group='WorldMap') %>% 
  addPolylines(data=streetNet2,group='All Links',color=~mpal2(ifelse(AB_FC==9,2,1)),weight = 2,
               label=lapply(streetNet2@data$tazLabel, function(x) HTML(x)),
                labelOptions = labelOptions(noHide = F,offset = c(10,-20))) %>% 
  addPolylines(data=streetNet2_wCounts,group='Links w Counts',weight = streetNet2_wCounts$thickCat,
               color=~mpal(colorCat),
               label=lapply(streetNet2_wCounts@data$tazLabel, function(x) HTML(x)),
                labelOptions = labelOptions(noHide = F,offset = c(10,-20))) %>%   
    addCircleMarkers(lng = streeCen$x, lat = streeCen$y,
                     label = as.character((streeCen$DAILYVOLUME)),group='Model Volume',opacity=0,radius = 1,
                     labelOptions = lapply(1:nrow(streeCen), function(x) {
                         labelOptions(opacity=1, noHide = T,textOnly = TRUE,textsize='10px',
                                      direction = 'auto', offset=c(0,0)) 
                       
                     })) %>%   
    addLayersControl(
        baseGroups = c('WorldGray','WorldMap'),
        overlayGroups = c('All Links','Links w Counts','Model Volume'),
        options = layersControlOptions(collapsed = F)  ) %>% 
    hideGroup(c('Model Volume'))
mm  

```


## Transit Validation

Transit assignment validation compares the observed transit boardings against the model assigned transit boardings. The table below shows the route level comparison and the overall comparison. The overall boardings are very close to the observed boardings; however, at route level there are substantial differences. Matching transit boardings at a route level is challenging problem in a regional model but can be improved by improving the transit network coding and making it consistent across the region, which will also involve updating the highway network. Walk access distance is a significant determining factor for transit use; making its representation more realistic and breaking down large TAZs along the transit use corridor will also help in modeling transit more accurately. 

```{r, include=T,messages=F,echo=F}



routeSumm <- rbind.data.frame(SummerLateNightboardAlight,SummerAMPeakboardAlight,SummerMiddayboardAlight,SummerPMPeakboardAlight) %>% 
  left_join(Tahoe_TransitR %>% select(ROUTE_ID,ROUTE_NAME),by=c('ROUTE'='ROUTE_ID')) %>% 
  left_join(transitRouteMap,by=c('ROUTE_NAME'='Route_Name_Dir')) %>% 
  group_by(`Route Desc`) %>% summarise(`Model Boardings`= sum(ALIGHTEND))
boardingSumm <- boardingTargets %>% full_join(routeSumm,by=c('Route Desc')) %>% replace_na(list(`Model Boardings`=0))

boardingSumm2 <- boardingSumm %>% select(-Category) %>% rbind.data.frame(boardingSumm %>% summarise_if(is.numeric,sum) %>% mutate(`Route Desc`='Total'))
boardingSumm2 <- boardingSumm2 %>% mutate_if(is.numeric,round)

boardingSumm2 %>% kable(caption = 'Table 5: Boardings by Route')    %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

boardingSumm3 <- boardingSumm %>% group_by(Category) %>% 
  summarise(`Target Boardings`=sum(`Target Boardings`),
            `Model Boardings` = sum(`Model Boardings`))

```


## Vehicles Miles Travelled

Vehicle miles travelled (VMT) is the total distance travelled  by all the vehicle trips in the model, it quantifies the amount of travel on a network. These trips are segmented by the following categories:

 + through trips referred to as External-External (EE) trips, 

 + internal-external and external-internal trips (trip origin or destination outside the modeling area) referred to as  IE/EI trips

 + and internal trips (both trip origin and destination inside the modeling area) – II trips
 
The VMT target was obtained from StreetLight Data. The target data for number of trips, VMT and the average trip distance for each of the trip categories can be seen in Table 6. The model summary is shown in Table 7. The average trip distance, the VMT and the number of trips match well for the II trips and EE trips. However, for the IE trips, the number of trips and VMT in the model output is marginally higher than the StreetLight Targets. The reason is that the highway count data suggested a higher volume (than Streetlight Data counts) along some of the external stations and the model was validated to those higher counts. During model validation, count data typically takes priority over other data sources such as VMT targets.

### VMT Report

```{r,echo=F,eval=T}

systemLevelTarget2 <- systemLevelTarget %>% mutate(`Total VMT`=round(`Total VMT`,0),`Mean Distance` = round(`Mean Distance`,1))
tripSumm <- trip_file %>% group_by(TripCat) %>% 
  summarise(`NumTrip` = n(), 
            `Total VMT`=round(sum(DistanceTotal,na.rm=T),0),
            `Mean Distance` = round(mean(DistanceTotal,na.rm=T),1))

tripSumm2 <- trip_file %>% filter(TripCat=='IE Trips')
tripSumm2 <- tripSumm2 %>% summarise(
            `Inside the model area`=round(sum(DistanceWithinRegion,na.rm=T),0),
            `Outside the model area`=round(sum(ExtDist,na.rm=T),0)) %>% gather(TripCat,'Total VMT') %>% 
  left_join(tripSumm2 %>% 
    summarise(
            `Inside the model area`=round(mean(DistanceWithinRegion,na.rm=T),1),
            `Outside the model area`=round(mean(ExtDist,na.rm=T),1)) %>% gather(TripCat,'Mean Distance'), by = c("TripCat")
  ) %>% mutate(NumTrip=NA)

tripSumm3 <- trip_file %>% filter(TripCat=='EE Trips') 
tripSumm3 <- tripSumm3 %>% summarise(
            `Inside the model area`=round(sum(DistanceWithinRegion,na.rm=T),0),
            `Outside the model area`=round(sum(ExtDist,na.rm=T),0)) %>% gather(TripCat,'Total VMT') %>% 
  left_join(tripSumm3 %>% 
    summarise(
            `Inside the model area`=round(mean(DistanceWithinRegion,na.rm=T),1),
            `Outside the model area`=round(mean(ExtDist,na.rm=T),1)) %>% gather(TripCat,'Mean Distance'), by = c("TripCat")
  ) %>% mutate(NumTrip=NA)


tripSummFnl <- rbind.data.frame(tripSumm[1,],tripSumm3,tripSumm[2,],tripSumm2,tripSumm[3,])

# Inside the model area
# Outside the model area



tempRowSum <- systemLevelTarget2 %>% filter(!is.na(`Num Trips`)) %>% summarise_if(is.numeric,sum) %>% 
  mutate(TripCat='Total',`Mean Distance`=`Total VMT`/`Num Trips`)
systemLevelTarget2 <- rbind.data.frame(systemLevelTarget2,tempRowSum) %>% mutate_at(4,round,1)

tempRowSum <- tripSummFnl %>% filter(!is.na(`NumTrip`)) %>% summarise_if(is.numeric,sum) %>% 
  mutate(TripCat='Total',`Mean Distance`=`Total VMT`/NumTrip)
tripSumm <- rbind.data.frame(tripSummFnl,tempRowSum) %>% mutate_at(4,round,1)

tripSumm$TripCat[tripSumm$TripCat=='IE Trips'] <- 'IE/EI Trips'

systemLevelTarget2 %>% 
  kable(caption='Table 6: Target Trip Totals,  VMT and Mean Distance',
        align=c('l',rep('r',3)),format.args = list(big.mark = ",")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))  %>% add_indent(c(3,4))


tripSumm %>%
  kable(caption='Table 7: Model Trip Totals, VMT and Mean Distance',
        align=c('l',rep('r',3)),format.args = list(big.mark = ",")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% add_indent(c(2,3,5,6))


```

Table 8 and Table 9 below shows the target and model summary for the EI (and IE) trips segmented by short-distance external zone trips and long-distance external zone trips (the definition and classification of trips into short-distance and long-distance is described in detail in the report *Create External Transportation Analysis Zones*). Number of trips and the VMT for each category is compared. It can be observed from the table that the model outputs match the target data reasonably well.

### Short and Long External Travel

```{r,echo=F,eval=T}

extTripsTarget <- extTripsTarget %>% left_join(extTripsVMTTarget,by='Gate') 
extTripsTarget %>% mutate_at(2:ncol(extTripsTarget),format,big.mark=',') %>% 
  kable(caption='Table 8: Target IE/EI Trips and VMT (Short vs. Long)',align='r') %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

tripSumm <- trip_file %>% filter(TripCat=='IE Trips') %>% 
  mutate(extTaz = ifelse(startTaz<=extTAZCutOff,startTaz,endTaz)) %>% filter(extTaz>0) %>% 
  group_by(extTaz) %>% 
  summarise(`NumTrip` = n(), 
            `Total VMT`=round(sum(DistanceTotal,na.rm=T),0))

tripSumm1 <- tripSumm %>% filter(extTaz<10)
names(tripSumm1) <- c('Gate','Short-Distance Trips','Short-Distance VMT')
tripSumm2 <- tripSumm %>% filter(extTaz>=10) %>% mutate(extTaz = extTaz/10)
names(tripSumm2) <- c('Gate','Long-Distance Trips','Long-Distance VMT')
tripSumm <- tripSumm1 %>% left_join(tripSumm2,by='Gate')

tempRowSum <- tripSumm %>% ungroup() %>% summarise_if(is.numeric,sum) %>% 
  mutate(Gate='Total')
tripSumm <- rbind.data.frame(tripSumm,tempRowSum)


tripSumm %>% 
   mutate_at(2:ncol(tripSumm),format,big.mark=',') %>% 
  kable(caption='Table 9: Model IE/EI Trips and VMT (Short vs. Long)',align='r') %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```