---
title: "Streetlight Data Summary for Lake Tahoe Region"
output:
  html_document:
    code_folding: hide
date: "January 2020"
---

```{r setup, include=FALSE,messages=F}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(htmltools)
```

To better validate the Tahoe Model outputs, Streetlight Data for the Tahoe region was gathered for the following time periods, including:

*  First 2 weeks in June 2018 (06/04 - 06/15)
*  Last week in August 2018 (08/27 - 08/31)
*  2nd and 3rd week in September 2018 (09/10 - 09/21)

For this study, trips on typical weekdays (Monday - Thursday) are analyzed, and trips on Fridays or weekends are not included. The [tidyverse](https://www.tidyverse.org/) package provides tools and functions used in this analysis. Please contact [WSP USA](mailto: R.Sidharthan@wsp.com, mingyang.li@wsp.com) for questions regarding the analysis. 


## Methodology 
### Gates
The objective of this analysis is to summarize the VMT from I-E and E-I trips as observed in the Streetlight Dataset. "Gates" are the major entrance / exit point to the Tahoe Lake region, and there are a total of 7 gates considered in this study. 

*  [Gate 1: Tahoe Meadows South Trailhead](https://www.google.com/maps/place/Tahoe+Meadows+South+Trailhead/@39.1601312,-120.1347057,10.4z/data=!4m5!3m4!1s0x80996e50dc563fbb:0x3dedbf0be21ccb41!8m2!3d39.3016431!4d-119.9186693)
*  [Gate 2: Spooner Summit Trailhead](https://www.google.com/maps/search/Spooner+Summit+Trailhead/@39.064722,-119.9944846,11.04z)
*  [Gate 3: Daggett Pass](https://www.google.com/maps/place/Daggett+Pass/@39.0958601,-120.0750635,11.04z/data=!4m5!3m4!1s0x80999a06e00dbe3d:0x7af0ad85570fe69b!8m2!3d38.9765746!4d-119.888791)
*  [Gate 4: Luther Pass](https://www.google.com/maps/place/Luther+Pass/@38.9902042,-120.1827023,10.16z/data=!4m5!3m4!1s0x8099ed3e3ced9ca7:0xb76eafccc32539f4!8m2!3d38.7868509!4d-119.9468476)
*  [Gate 5: Echo Summit](https://www.google.com/maps/place/Echo+Summit/@39.0275357,-120.4109167,10.26z/data=!4m5!3m4!1s0x80998cfa5eee2bb9:0x489132d7a3ea7f76!8m2!3d38.8129628!4d-120.0301833)
*  [Gate 6: Truckee River Bike Trail](https://www.google.com/maps/place/Truckee+River+Bike+Trail,+California/@39.0505355,-120.4193488,10.02z/data=!4m5!3m4!1s0x809bd7eec370906d:0xb4f212616f01994!8m2!3d39.1773546!4d-120.1906817)
*  [Gate 7: N Shore Blvd](https://www.google.com/maps/place/N+Shore+Blvd,+NORTHSTAR,+CA+96161/@39.1221507,-120.3170886,10.28z/data=!4m5!3m4!1s0x8099638234461bdd:0x6703827436e94fb6!8m2!3d39.2583015!4d-120.0667665)

### Assumptions
Once the gates are defined, the study team make the following assumptions regarding the travel patterns.

+  There are 8 internal zones for the Tahoe region. The block groups outside of the Tahoe region are external zones. 
+  All external zones can be grouped in 1 of 29 external districts. 
+  All I-E and E-I trips between a external district and a internal zone will use the same gate.
+  The Streetlight data contains the Annual Average Daily Traffic (AADT) for I-E and E-I trips
+  The I-E and E-I trips contains 2 parts:
    +  distance between external zone and gate (external_distance), and
    +  distance between gate and internal zone (internal_distance), which should be the same for all the trips between the same gate and same internal zone
+  external_distance will be used to further categorize the I-E and E-I trips:

    +  trips with external_distance <= 60 miles are considered **short-distance trips**
    +  trips with external_distance > 60 miles are considered **long-distance trips**

### VMT and Trip Summaries      
Based on the inputs and assumptions, VMT can be calculated by multiplying AADT with external_distance and internal_distance. The results are then summarized by gates and by trip types (long- vs. short-distance trips). The summary tables will be used to validate the model outputs. 

+  VMT (Outside TRPA Region)
+  VMT (Within TRPA Region)
+  Total Number of Trips


## Streetlight Data Overview
There are 3 Streetlight databases, a snapshot of the datasets are found below: 

```{r, message=F, warning=F, results = 'asis', echo = F}
dataset_1 <- read_csv('../Data/95945_Tahoe_OD_Analysis_2020_01_02_odg_all.csv')
kable(dataset_1[1:3, 2:6], align = c(rep('c',5)), caption = "1. Trip information dataset") %>%
    kable_styling()

dataset_2 <- read_csv('../Data/95945_Tahoe_OD_Analysis_2020_01_02_odg_traveler_all.csv')
kable(dataset_2[1:3, 30:34], align = c(rep('c',5)), caption = "2. Traveller information dataset") %>%
    kable_styling()

dataset_3 <- read_csv('../Data/95945_Tahoe_OD_Analysis_2020_01_02_odg_trip_all.csv')
kable(dataset_3[1:3, 23:27], align = c(rep('c',5)), caption = "3. Additional trip information dataset") %>%
    kable_styling()
```


## Data Processing and Summary
There are 5 major steps during this phase, including:

1.  Import Streetlight data
```{r,message=F,eval=F,warning=F}
# import data from local folder, and select the rows that are not all-day redcords and are on typical weekdays 
dataset_1 <- read_csv('../Data/95945_Tahoe_OD_Analysis_2020_01_02_odg_all.csv',guess_max=10000000) %>% 
    data.frame() %>% as_tibble() %>%
    filter(Day.Part!='0: All Day (12am-12am)') %>% 
    filter(Day.Type=='1: Weekday (M-Th)') 

dataset_2 <- read_csv('../Data/95945_Tahoe_OD_Analysis_2020_01_02_odg_traveler_all.csv',guess_max=10000000)%>%
    data.frame() %>% as_tibble() %>% 
    filter(Day.Part!='0: All Day (12am-12am)') %>% 
    filter(Day.Type=='1: Weekday (M-Th)') 

dataset_3 <- read_csv('../Data/95945_Tahoe_OD_Analysis_2020_01_02_odg_trip_all.csv',guess_max=10000000) %>%
    data.frame() %>% as_tibble() %>% 
    filter(Day.Part!='0: All Day (12am-12am)') %>% 
    filter(Day.Type=='1: Weekday (M-Th)') 
```

2.  Trim the Streetlight data and combine the data for further analysis
```{r,message=F,eval=F,warning=F}
# trim the undesired attributes in the data
dataset_3 <- dataset_3 %>% 
    select(Origin.Zone.Name, Destination.Zone.Name, Day.Part,Avg.All.Trip.Length..mi.,Avg.Trip.Length..mi.)
dataset_2 <- dataset_2 %>%
    select(Day.Part,Origin.Zone.Name,Destination.Zone.Name,Purpose.HBW..percent.,Purpose.HBO..percent.,
           Purpose.NHB..percent.)

# combine the Streetlight data, calculate the home-based work(HBW), home-based other(HBO), and non-home-based(NHB) trips, and trim attributes that are not needed
data_v1 <- dataset_1 %>% 
    left_join(dataset_2) %>% 
    left_join(dataset_3) %>% 
    mutate(Purpose.HBW..percent.=as.numeric(Purpose.HBW..percent.),
           Purpose.HBO..percent.=as.numeric(Purpose.HBO..percent.),
           Purpose.NHB..percent.=as.numeric(Purpose.NHB..percent.)) %>% 
    replace_na(list(Purpose.HBW..percent.=0,Purpose.HBO..percent.=0,Purpose.NHB..percent.=0)) %>% 
    mutate(HBW_vol = Average.Daily.O.D.Traffic..StL.Volume.*Purpose.HBW..percent.,
           HBO_vol = Average.Daily.O.D.Traffic..StL.Volume.*Purpose.HBO..percent.,
           NHB_vol = Average.Daily.O.D.Traffic..StL.Volume.*Purpose.NHB..percent.,
           Tot_vol = Average.Daily.O.D.Traffic..StL.Volume.) %>% 
    select(-Type.of.Travel,-Day.Type,
           -Origin.Zone.Is.Pass.Through,-Origin.Zone.Direction..degrees.,-Origin.Zone.is.Bi.Direction,
           -Destination.Zone.Is.Pass.Through,-Destination.Zone.Direction..degrees.,-Destination.Zone.is.Bi.Direction,
           -Avg.Trip.Duration..sec.,-Origin.Zone.ID,-Destination.Zone.ID,
           -Average.Daily.Origin.Zone.Traffic..StL.Volume.,-Average.Daily.Destination.Zone.Traffic..StL.Volume.,
           -Purpose.HBW..percent.,-Purpose.HBO..percent.,-Purpose.NHB..percent.,
           -Average.Daily.O.D.Traffic..StL.Volume.)
```

3.  Import 3 additional input files with gate and external district information. 

```{r, message=F, warning=F, results=F}
# define the relationship between the block groups and the 29 external districts
CA_NV_tract_to_district <- read_csv('../Data/Inputs/CA_NV_Tract_To_District.csv')
kable(CA_NV_tract_to_district[1:3,], align = c(rep('c',ncol(CA_NV_tract_to_district))), caption = "External districts information") %>%
    kable_styling()

# define which gate to use for E-I and I-E trips. Gates are determined based on Gogole Map routing information. Assumed departure time on Tuesday, 7:30 AM 
district_to_gate_relationship <- read_csv('../Data/Inputs/District_to_Gate.csv') %>%
    gather(Destination, Gate, "06_0_Chico":"Grand Total") %>%
    filter(!is.na(Gate)) %>%
    rename("Origin" = "Row Labels")
kable(district_to_gate_relationship[1:3,], align = c(rep('c',ncol(district_to_gate_relationship))), caption = "Gate information") %>%
    kable_styling()

# define distance between gate and internal zones. Distance measured from gate location to the population centroid of the internal zones
gate_to_internal_distance <- read_csv('../Data/Inputs/Gate_to_Internal.csv') %>%
    gather(Internal, Distance, "11":"18")
kable(gate_to_internal_distance[1:3,], align = c(rep('c',ncol(gate_to_internal_distance))), caption = "Gate to internal zone distance") %>%
    kable_styling()
```

4. Analyze the combined dataset to further evaluate the O-D trip patterns by incorporating the inputs from Step 3.

```{r message=FALSE, warning=FALSE, eval=F}
# add the state, county, and tract information, and retain only I-E and E-I trips
data_v2 <- data_v1 %>% 
    mutate(orig_state=ifelse(Origin.Zone.Source!='Input',substr(Origin.Zone.Name,2,3),'Input'),
           orig_county=ifelse(Origin.Zone.Source!='Input',substr(Origin.Zone.Name,4,6),'Input'),
           orig_tract=ifelse(Origin.Zone.Source!='Input',substr(Origin.Zone.Name,7,12),Origin.Zone.Name),
           dest_state=ifelse(Destination.Zone.Source!='Input',substr(Destination.Zone.Name,2,3),'Input'),
           dest_county=ifelse(Destination.Zone.Source!='Input',substr(Destination.Zone.Name,4,6),'Input'),
           dest_tract=ifelse(Destination.Zone.Source!='Input',substr(Destination.Zone.Name,7,12),Destination.Zone.Name)) %>%
    left_join(CA_NV_tract_to_district %>%
                  rename(orig_district=groupname),by=c('orig_state'='STATEFP','orig_county'='COUNTYFP',
                                                      'orig_tract'='TRACTCE')) %>% 
    left_join(CA_NV_tract_to_district %>%
                  rename(dest_district=groupname),by=c('dest_state'='STATEFP','dest_county'='COUNTYFP',
                                                      'dest_tract'='TRACTCE')) %>% 
    mutate(orig_district=ifelse(Origin.Zone.Source=='Input', Origin.Zone.Name, orig_district)) %>% 
    mutate(dest_district=ifelse(Destination.Zone.Source=='Input', Destination.Zone.Name, dest_district)) %>% 
    filter(!(is.na(orig_district)|is.na(dest_district)))

# calculate the statistics of the trip length, trip volumes, by oirigin and destination.
data_v3 <- data_v2 %>% 
    mutate(Avg.All.Trip.Length..mi.=as.numeric(Avg.All.Trip.Length..mi.),
           Avg.Trip.Length..mi.=as.numeric(Avg.Trip.Length..mi.),
           ) %>% 
    mutate(Avg.Trip.Length..mi.=ifelse(is.na(Avg.Trip.Length..mi.),0,Avg.Trip.Length..mi.)) %>% 
    group_by(Origin.Zone.Name,Destination.Zone.Name) %>% 
    summarize( orig_state=first(orig_state),
               orig_county=first(orig_county),
               orig_tract=first(orig_tract),
               dest_state=first(dest_state),
               dest_county=first(dest_county),
               dest_tract=first(dest_tract),
               orig_district=first(orig_district),
               dest_district=first(dest_district),
               mean_distance = weighted.mean(Avg.All.Trip.Length..mi.,Tot_vol),
               HBW_vol=sum(HBW_vol),
               HBO_vol=sum(HBO_vol),
               NHB_vol=sum(NHB_vol),
               Tot_vol=sum(Tot_vol),
               ) %>%
    ungroup() %>%
    filter(!is.na(mean_distance))

# set the threshold for long-distance and short-distance trips
long_distance_threshold = 60

# define the internal and external zones for the I-E and E-I trips, and categorize the trips based on the average external_distance
# update the internal and external distances if the calculated external distance is negative
data_v4 <- data_v3 %>%
    left_join(district_to_gate_relationship, by=c('orig_district'='Origin','dest_district'='Destination')) %>%
    filter(!is.na(Gate)) %>% # Remove internal-internal trips
    rename(gate = Gate) %>%
    mutate(internal_zone = ifelse(orig_district %in% c('11','12','13','14','15','16','17','18'),
                                 orig_district, dest_district),
           external_zone = ifelse(Origin.Zone.Name %in% c('11','12','13','14','15','16','17','18'),
                                  Destination.Zone.Name, Origin.Zone.Name)) %>%
    left_join(gate_to_internal_distance, by=c('gate'='gate','internal_zone'='Internal')) %>%
    rename(internal_distance = Distance) %>%
    mutate(external_distance_to_gate = mean_distance - internal_distance,
           flag = ifelse(external_distance_to_gate < 0, 1, 0), 
           long_distance = ifelse(external_distance_to_gate > long_distance_threshold, 1, 0)) %>%
    mutate(external_distance_to_gate = ifelse(flag == 1, 0, external_distance_to_gate),
           internal_distance = ifelse(flag == 1, mean_distance, internal_distance))

```

5. Summarize the VMT and trips by gates and trip types

```{r,message=F, warning=F, eval=F}
# compute desired attributes
data_v5 <- data_v4 %>%
    group_by(internal_zone, gate) %>%
    summarize(total_trip = sum(Tot_vol),
              external_VMT = sum(Tot_vol*external_distance_to_gate),
              internal_VMT = sum(Tot_vol*internal_distance),
              long_distance_external_VMT = sum(ifelse(long_distance == 1, Tot_vol*external_distance_to_gate, 0)),
              short_distance_external_VMT = sum(ifelse(long_distance == 1, 0, Tot_vol*external_distance_to_gate)),
              long_distance_internal_VMT = sum(ifelse(long_distance == 1, Tot_vol*internal_distance, 0)),
              short_distance_internal_VMT = sum(ifelse(long_distance == 1, 0, Tot_vol*internal_distance)),
              long_distance_trip = sum(ifelse(long_distance == 1, Tot_vol, 0)),
              short_distance_trip = sum(ifelse(long_distance == 1, 0, Tot_vol)))
    
# define a function to output the data by gates and internal zones
summarize_data_1 <- function(sum_attribute){
    data_temp <- data_v5 %>% 
        select(internal_zone, gate, sum_attribute) %>%
        spread(internal_zone, sum_attribute) %>%
        replace_na(list("11"=0,"12"=0,"13"=0,"14"=0,"15"=0,"16"=0,"17"=0,"18"=0)) %>%
        rename("Gate / Internal zones" = "gate")
    write_csv(data_temp, paste0(sum_attribute,"_summary_1.csv"))
}

# call the function
summarize_data_1("external_VMT")
summarize_data_1("internal_VMT")
summarize_data_1("long_distance_external_VMT")
summarize_data_1("short_distance_external_VMT")
summarize_data_1("long_distance_internal_VMT")
summarize_data_1("short_distance_internal_VMT")
summarize_data_1("total_trip")
summarize_data_1("long_distance_trip")
summarize_data_1("short_distance_trip")

# group desired summaries by gates
data_v6 <- data_v5 %>%
    group_by(gate) %>%
    summarize(short_distance_external_VMT = sum(short_distance_external_VMT) %>% round(),
              long_distance_external_VMT = sum(long_distance_external_VMT) %>% round(),
              short_distance_internal_VMT = sum(short_distance_internal_VMT) %>% round(),
              long_distance_internal_VMT = sum(long_distance_internal_VMT) %>% round(),
              short_distance_trip = sum(short_distance_trip) %>% round(),
              long_distance_trip = sum(long_distance_trip) %>% round())
    
# define a function to output the srummaries by gates
summarize_data_2 <- function(attribute_1, attribute_2){
    data_temp <- data_v6 %>% 
        select(gate, attribute_1, attribute_2) %>%
        rename("Gate" = "gate",
               "Short-Distance Trips" = paste0(attribute_1),
               "Long-Distance Trips" = paste0(attribute_2))
    write_csv(data_temp, paste0(attribute_1, "_", attribute_2, "_summary_2.csv"))
}


# call the function
summarize_data_2("short_distance_external_VMT", "long_distance_external_VMT")
summarize_data_2("short_distance_internal_VMT", "long_distance_internal_VMT")
summarize_data_2("short_distance_trip", "long_distance_trip")

```

```{r,message=F, warning=F, echo = F, result = "asis"}
#Show the summaries
data_temp <- read_csv("short_distance_external_VMT_long_distance_external_VMT_summary_2.csv")
kable(data_temp, format.args = list(big.mark = ","), align = c(rep('c',ncol(data_temp))), caption = "VMT (Outside TRPA Region)") %>%
    kable_styling() 
data_temp <- read_csv("short_distance_internal_VMT_long_distance_internal_VMT_summary_2.csv")
kable(data_temp, format.args = list(big.mark = ","), align = c(rep('c',ncol(data_temp))), caption = "VMT (Within TRPA Region)") %>%
    kable_styling() 
data_temp <- read_csv("short_distance_trip_long_distance_trip_summary_2.csv")
kable(data_temp, format.args = list(big.mark = ","), align = c(rep('c',ncol(data_temp))), caption = "Total Number of Trips") %>%
    kable_styling() 
```

```{r,output=F,message=F,echo=F,eval=F,results=F}
#categorize the block group to be either "long-distance" or "short-distance", and to rename them by gate number
#short-distance zones are named 1 - 7, long-distance zones are named 10, 20, 30, ..., 70
#this is strictly for mapping purposes
data_v6 <- data_v4 %>%
    group_by(external_zone, gate) %>%
    summarize(total_trip = sum(Tot_vol),
              long_distance_per = sum(Tot_vol*long_distance)/total_trip * 100,
              bg_type = ifelse(long_distance_per >= 50, "long-distance", "short-distance"))  %>%
    ungroup() %>%
    group_by(external_zone) %>%
    filter(total_trip == max(total_trip)) %>%
    ungroup() %>%
    mutate(external_zone = str_sub(external_zone,2,-2),
           zone_group = ifelse(bg_type == "short-distance", gate, paste0(gate,"0")))

write_csv(data_v6, "summary_by_bg.txt")
```